---
title: "ASX - Machine Learning"
author: "Ben Cole - s3412349"
date: "Print Date: `r format(Sys.Date(), '%d/%m/%Y')`"
output: html_document
---

# **Phase 1 - Introduction and Cleaning**

## Outline

The aim of this supervised machine learning project is to predict the prices of 50 different Australian Stock Exchange (ASX) shares in the year 2019. Data beginning at the 2019 calendar year through to April 2019 was used in the training and testing data sets. 

### Data

The data used was historical summary data of all shares available with a trading history in the ASX between 02/01/2019 through to business week (Mon - Fri) ending 12/04/2019. The data was provided by the website [ASX Historical Data](!"https://www.asxhistoricaldata.com/"). The data was compressed into .zip files separated by calendar month between 02/01/2019 - 31/01/2019 and then by business week from 01/02/2019 - 12/04/2019. The raw data followed the same structure throughout all text files, and was not provided with headers. Each comma separated value followed the following headers:

* Ticker - the three-digit unique identifier ASX ticker code (renamed to `ASX_Ticker`)
* Date - date of trade information
* Open - price per individual share at the beginning of the day's trade
* High - highest price recorded per individual share during the day's trade
* Low - lowest price recorded per individual share during the day's trade
* Close - price per individual share at the end of the day's trade
* Volume - number of shares traded during the day


### Packages

```{r}

library(pacman) ## for loading multiple packages

suppressMessages(p_load(character.only = T,
                        install = F,
                        c("tidyverse",  ## thanks Hadley
                          "lubridate",  ## for handling dates
                          "zoo",        ##
                          "lemon",      ## add ons for ggplot
                          "knitr",      ## knitting to RMarkdown
                          "kableExtra", ## add ons for knitr tables
                          "scales",     ## quick and easy formatting prettynums
                          "e1071")))    ##

```

### Data

The data was read making use of a nested for loop for the files that were separated by week. Just a single for loop was required for the data that was collated into January 2019.

```{r}

Jan_file <- list.files(pattern = "jan")

unzip(Jan_file)

Jan_File_no_zip <- list.files(pattern = "jan")[!str_detect(list.files(pattern = "jan"), ".zip")]

ASX_Data_Week_Jan <- list()

ASX_Data_Month_Jan <- list()

for (k in 1:length(list.files(Jan_File_no_zip))) {
  
  ASX_Data_Week_Jan[[k]] <- read_csv( file.path(Jan_File_no_zip,
                                                   list.files(Jan_File_no_zip)[k]),
                                           col_names = c("ASX_Ticker",
                                                         "Date",
                                                         "Open",
                                                         "High",
                                                         "Low",
                                                         "Close",
                                                         "Volume") )
  
  ASX_Data_Month_Jan[[k]] <- do.call(rbind, ASX_Data_Week_Jan)
  
}


h <- 1

repeat {
  
  unzip(list.files(pattern = "week")[h])
  
  h <- h+1
  
  if (h > length(list.files(pattern = "week"))) {
    break
  }
  
}

Week_files <- list.files(pattern = "week")
Zip_files <- list.files(pattern = ".zip")

Week_files_no_zip <- Week_files[!Week_files %in% Zip_files]

ASX_Data_List <- list()

ASX_Data_List_Week <- list()

for (i in 1:length(Week_files_no_zip)){
  
  for (j in 1:length(list.files(path=Week_files_no_zip[i]))){
    
    ASX_Data_List_Week[[j]] <- read_csv(file.path(Week_files_no_zip[i],
                                               list.files(Week_files_no_zip[i])[j]),
                                        col_names=c("ASX_Ticker",
                                                    "Date",
                                                    "Open",
                                                    "High",
                                                    "Low",
                                                    "Close",
                                                    "Volume"))
  }
  
  ASX_Data_List[[i]] <- do.call(rbind, ASX_Data_List_Week)
  
}


ASX_Data_Frame_Jan <- do.call(rbind, ASX_Data_Month_Jan)

ASX_Data_Frame_Post_Jan <- do.call(rbind, ASX_Data_List)

ASX_Data_Frame <- rbind(ASX_Data_Frame_Jan,
                        ASX_Data_Frame_Post_Jan)

max(ASX_Data_Frame$Date)

head(ASX_Data_Frame, 20)

sample_n(ASX_Data_Frame, size=20)

# rm(ASX_Data_List, ASX_Data_List_Temp)

```


### Descriptive Statistics

```{r}

ASX_Long <- gather(ASX_Data_Frame,
                   Open:Volume,
                   key="Variable",
                   value="Value")

ASX_Summary <- summarise(group_by(ASX_Long,
                                  Variable),
                         "n ASX_Tickers" = comma(length(unique(ASX_Ticker))),
                         "n Observations" = comma(n()),
                         "Min Date" = format(ymd(min(Date)), "%d/%m/%Y"),
                         "Max Date" = format(ymd(max(Date)), "%d/%m/%Y"),
                         "Minimum" = comma(min(Value)),
                         "Q1" = comma(quantile(Value, 0.25)),
                         "Median" = comma(quantile(Value, 0.5)),
                         "Q3" = comma(quantile(Value, 0.75)),
                         "90th Percentile" = comma(quantile(Value, 0.9)),
                         "95th Percentile" = comma(quantile(Value, 0.95)),
                         "Maximum" = comma(max(Value)),
                         "Skew" = round(skewness(Value), 3),
                         "Kurtosis" = round(kurtosis(Value), 2),
                         "NA count" = comma(sum(is.na(ASX_Data_Frame))))

kable_styling(kable(ASX_Summary),
              full_width = T,
              bootstrap_options = c("striped"),
              position = "float_left")

```


### Plots

```{r}

ggplot(ASX_Long) +
  geom_density(aes(x=Value)) +
  scale_x_continuous(labels=comma) +
  facet_rep_wrap(~Variable) +
  theme_minimal()

```


### Filtering data

```{r}

ASX_Ticker_Summary <- summarise(group_by(ASX_Data_Frame, ASX_Ticker),
                                "n ASX_Tickers" = comma(length(unique(ASX_Ticker))),
                                "n Observations" = comma(n()),
                                "Min Date" = format(ymd(min(Date)), "%d/%m/%Y"),
                                "Max Date" = format(ymd(max(Date)), "%d/%m/%Y"),
                                "Minimum" = min(High),
                                "Q1" = quantile(High, 0.25),
                                "Median" = quantile(High, 0.5),
                                "Q3" = quantile(High, 0.75),
                                "90th Percentile" = quantile(High, 0.9),
                                "95th Percentile" = quantile(High, 0.95),
                                "Maximum" = max(High),
                                "Skew" = round(skewness(High), 3),
                                "Kurtosis" = round(kurtosis(High), 2),
                                "NA count" = comma(sum(is.na(ASX_Data_Frame))))
  

kable(sample_n(ASX_Ticker_Summary, 20))

ASX_Lower <- filter(ASX_Ticker_Summary, Maximum < quantile(Maximum, 2/3))
  
nrow(ASX_Lower)

ASX_Long_Lower <- filter(ASX_Long, ASX_Ticker %in% ASX_Lower$ASX_Ticker)

```


### Removing Extreme ASX_Tickers

```{r}

ASX_Summary_Lower <- summarise(group_by(ASX_Long_Lower,
                                        Variable),
                               "n ASX_Tickers" = comma(length(unique(ASX_Ticker))),
                               "n Observations" = comma(n()),
                               "Min Date" = format(ymd(min(Date)), "%d/%m/%Y"),
                               "Max Date" = format(ymd(max(Date)), "%d/%m/%Y"),
                               "Minimum" = round(min(Value), 2),
                               "Q1" = round(quantile(Value, 0.25), 2),
                               "Median" = round(quantile(Value, 0.5), 2),
                               "Q3" = round(quantile(Value, 0.75), 2),
                               "90th Percentile" = round(quantile(Value, 0.9), 2),
                               "95th Percentile" = round(quantile(Value, 0.95), 2),
                               "Maximum" = round(max(Value), 2),
                               "Skew" = round(skewness(Value), 3),
                               "Kurtosis" = round(kurtosis(Value), 2),
                               "NA count" = comma(sum(is.na(ASX_Data_Frame))))

kable_styling(kable(ASX_Summary_Lower),
              bootstrap_options = c("striped"),
              position = "float_left",
              full_width = F)

```

